---
title: "Exporatory Data Analysis"
author: "Talha S."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(lubridate)
library(plotly)
library(gridExtra)
library(gganimate)
theme_set(theme_light())
```

```{r import, message=FALSE}
movie_profit <- read_csv("data/movie_profit.csv")
```

## Release Timing

### What are the most popular release dates?

```{r release-dates}
movie_profit %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(10) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top 10 most popular release dates",
    x = "",
    y = ""
  )
```

Christmas stands out as the most popular time for movie releases.

### Do movies of different genres get released on different dates of the year?

```{r release-dates-by-genre}

grid.arrange(
  movie_profit %>% 
  filter(genre=="Action") %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(4) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top Action movies release dates",
    x = "",
    y = ""
  ),
  movie_profit %>% 
  filter(genre=="Adventure") %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(4) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top Adventure movies release dates",
    x = "",
    y = ""
  ),
  movie_profit %>% 
  filter(genre=="Horror") %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(4) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top Horror movies release dates",
    x = "",
    y = ""
  ),
  movie_profit %>% 
  filter(genre=="Drama") %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(4) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top Drama movies release dates",
    x = "",
    y = ""
  ),
  movie_profit %>% 
  filter(genre=="Comedy") %>% 
  group_by(md, month_day) %>% 
  tally(sort = TRUE) %>% 
  head(4) %>% 
  ggplot(aes(fct_reorder(month_day, md), n, fill = month_day)) + 
  geom_col() + theme(legend.position = "none") +
  labs(
    title = "Top Comedy movies release dates",
    x = "",
    y = ""
  )
)
```

Horror Movie genre stands out in this regard as movies of all other genres come out the most on Christmas day, but halloween season suits the Horror genre.

## Production Budget

### Which distributors have poured in the most budget per movie?

```{r budget-per-distributor}
movie_profit %>% 
  group_by(distributor_top7) %>% 
  summarise(budget_per_movie = mean(production_budget)) %>%
  ggplot(aes(fct_reorder(distributor_top7, budget_per_movie), budget_per_movie, fill = distributor_top7)) +
  geom_col() + 
  coord_flip() + 
  labs(
    title = "Average budget per movie",
    x = "Distributor",
    y = "Budget per Movie"
  )
```

### Hollywood's best pay offs

```{r best-pay-offs}
movie_profit %>%
  mutate(profit = worldwide_gross - production_budget) %>% 
  arrange(desc(profit)) %>%
  head(15) %>% 
  ggplot(aes(y=profit, x=fct_reorder(movie, profit), fill = mpaa_rating)) + 
  geom_col() + 
  coord_flip() + 
  labs(
    title = "Most Profitable movies",
    x = "",
    y = "Net Profit"
  )
```

```{r best-and-worst}
movie_profit %>% 
  group_by(genre) %>% 
  summarise(Best = max(profit_ratio),
            Worst = min(profit_ratio))
```

### How rewarding are high production budgets?

```{r ROI}
movie_profit %>% 
  ggplot(aes(production_budget, worldwide_gross)) + 
  geom_point(aes(colour = genre), alpha = 0.2) + 
  scale_x_log10(labels = dollar_format()) + scale_y_log10(labels = dollar_format()) +
  geom_smooth(method = "lm", se = FALSE, colour = "red") + 
  labs(
    title = "Movie Return on Investment",
    x = "Production Budget (USD)",
    y = "Worldwide Gross (USD)"
  )
```

## Revenue & Budget: Last 50 years of cinema

```{r revenue-budget-50-years}
movie_profit %>% 
  filter(year(release_date) > 1968,
         distributor_top7 != "Other" & distributor_top7 != "Lionsgate") %>% 
  group_by(distributor_top7, decade) %>% 
  summarise(
    production_budget = sum(production_budget),
    worldwide_gross = sum(worldwide_gross),
    movies = n()
  ) %>% 
  gather(key = RevenueBudget, value = Amount, -distributor_top7, -decade, -movies) %>% 
  ggplot(aes(x = distributor_top7, y = Amount, fill = RevenueBudget)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_y_continuous(labels = dollar_format()) + coord_flip() + 
  theme(legend.position = "none") +
# Add time transition to above plot
  labs(title = '50 Years of The Five Biggest Movie Distributors',
       subtitle = 'Worldwide Gross (Teal) & Production Budgets (Pink) Year: {frame_time}', x = "",
        y = 'U.S. Dollars') +
  transition_time(as.integer(decade)) +
  ease_aes('linear')
```



## Appendix


```{r exit}
knitr::knit_exit()
```

## Scrap Work


```{r scrap-revenue-budget-50-years}
movie_profit %>% 
  filter(year(release_date) > 1967) %>% 
  group_by(genre, decade) %>% 
  summarise(
    production_budget = median(production_budget),
    worldwide_gross = median(worldwide_gross),
    movies = n()
  ) %>% 
  ggplot(aes(production_budget, worldwide_gross, colour = genre)) +
  geom_point(aes(size = movies)) +
  scale_x_log10(labels = dollar_format()) + 
  scale_y_log10(labels = dollar_format()) + 
# Add time transition to above plot
  labs(title = 'Revenue & Budget: Last 50 years of cinema',
       subtitle = 'Year: {frame_time}', 
       x = 'Median Production Budget', y = 'Median Worldwide Gross Revenue') +
  transition_time(as.integer(decade)) +
  ease_aes('linear')
```
